---
title: 'DATA 605:  Week 1 Assignment'
author: "Aaron Grzasko"
date: "February 5, 2017"
output: 
    html_document:
        theme: default
        highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem Set 1  
  
You can think of vectors representing many dimensions of related information. For
instance, Netflix might store all the ratings a user gives to movies in a vector. This is
clearly a vector of very large dimensions (in the millions) and very sparse as the user might
have rated only a few movies. Similarly, Amazon might store the items purchased by a user
in a vector, with each slot or dimension representing a unique product and the value of the
slot, the number of such items the user bought. One task that is frequently done in these
settings is to find similarities between users. And, we can use dot-product between vectors
to do just that. As you know, the dot-product is proportional to the length of two vectors
and to the angle between them. In fact, the dot-product between two vectors, normalized
by their lengths is called as the cosine distance and is frequently used in recommendation
engines
  
**1. Calculate the dot product u.v where u = [0.5; 0.5] and v = [3; -4]**  
  
*Given:*  
```{r}
u <- c(0.5, 0.5)
v <- c(3, -4)  

```
  
Below is a function to calculate the dot product of two vectors.  The function includes error handling procedures. 
  
```{r}
dotprod <- function(a,b){
    # check for numeric input
    if (!(is.numeric(a) && is.numeric(b))){
    warning("Both inputs must be numeric")
    }
    # verify that inputs are vector types or matrices with dimensions consistent with vector types
    else if (!(is.vector(a) || (is.matrix(a) && nrow(a) == 1) || (is.matrix(a) && ncol(a) == 1)) || 
             !(is.vector(b) || (is.matrix(b) && nrow(b) == 1) || (is.matrix(b) && ncol(b) == 1)))
        {
        warning("Inputs must be vectors, matrices with one row, or matrices with one column")
    } 
    # verify that vectors are of equal length
    else if (length(a) != length(b)){
        warning("The inputs must have equal lengths")
    }
    
    # calculate dot product; convert to vector form to ensure that function works appropriately for matrix inputs 
    else{
        sum(as.vector(a)*as.vector(b))
    }
}


```  
  
Here is the dot product of vectors u and v using dotprod():  
  
```{r}
dotprod(u,v)
```
  
As an alternative solution, I generated a dot product function using a for-loop:  
  
```{r}
# alternative homegrown function using loops
dotprod_alt <- function(a,b){
    
    if (!(is.numeric(a) && is.numeric(b))){
        warning("Both inputs must be numeric")
    }
    # verify that inputs are vector types or matrices with dimensions consistent with vector types
    else if (!(is.vector(a) || (is.matrix(a) && nrow(a) == 1) || (is.matrix(a) && ncol(a) == 1)) || 
             !(is.vector(b) || (is.matrix(b) && nrow(b) == 1) || (is.matrix(b) && ncol(b) == 1)))
    {
        warning("Inputs must be vectors, matrices with one row, or matrices with one column")
    } 
    # verify that vectors are of equal length
    else if (length(a) != length(b)){
        warning("The inputs must have equal lengths")
    }
    else {
        mydot <- 0
        
        for (i in 1:length(a)){
            mydot <- mydot +  a[i]*b[i]
        }
    }
    return(mydot)
}


```

  
Here is the dot product of u and v using the alternative, `dotprod_alt()` function:  
  
```{r}
dotprod_alt(u,v)

```
  
Finally, let's calculate the dot product of u and v using the built-in `crossprod()` function:  
  
```{r}
crossprod(u,v)

```  

**2. What are the lengths of u and v?**  
  
The easiest way to tacle this question is to use the dotprod() function created in the previous problem:  

```{r}
# Homegrown function to return length, i.e. norm2, of a vector, using custom dotprod() function
norm2 <- function(a){
    sqrt(dotprod(a,a))
}

``` 
  
Here are the requested lengths:  
  
```{r}
norm2(u) 
norm2(v)
```  
  
However, there are alternative ways to calculate the vector lengths:  
  
```{r}
# alternative norm2 function using a for loop
norm2_alt <- function(a){
    if (!is.numeric(a)){
        warning("input must be numeric")
    }
    else{
        mylength <- 0
        for (elem in a){
            mylength <- mylength + elem^2 
            
        }
    }
    mylength <- mylength^(0.5)
    return(mylength)
        
}
 

# another norm2 function without a loop, and also without referencing the previously created dotprod() function  
norm2_alt2 <- function(a){
    if (!(is.numeric(a))){ 
        warning("input must be numeric")
    }
    else{
        sqrt(sum(a^2))
    }
}

```
  
Let's check the solutions using these two, alternative functions:  
  
```{r}
norm2_alt(u)
norm2_alt(v)


norm2_alt2(u)
norm2_alt2(v)


```
  
Finally, we can calculate the lengths using the built in `norm()` function:  
  
```{r}
norm(u, type="2")
norm(v, type="2")
```
  
**3.  What is the linear combination of 3u - 2v?**  
  
R can easily handle this arithmetic without explicit calls to built-in or custom functions:  

```{r}
3*u - 2*v 

```
  
**4. What is the angle between u and v?**  
  
Below is a function to calculate the angle, in either radians or degrees.  I used the previously created functions, `dotprod()` and `norm2()` to produce the desired calculations.  
  
```{r}
angle_vec <- function(a,b,ang_type = 'rad'){
    acos(dotprod(a,b) / (norm2(a)*norm2(b))) * 
        if(ang_type == 'rad') 1 
        else if(ang_type == 'deg') 180/(2*pi)
        else warning('ang_type must be "rad" or "deg"')
} 


```

The angle between u and v is:  
```{r}
# radian measure
angle_vec(u,v)


# degree measure
angle_vec(u,v, "deg")
```
  
## Problem Set 2  
  
Set up a system of equations with 3 variables and 3 constraints and solve for x. Please write a function in R that will take two variables (matrix A & constraint vector b) and solve using elimination. Your function should produce the right answer for the system of equations for any 3-variable, 3-equation system. You don't have to worry about degenerate cases and can safely assume that the function will only be tested with a system of equations that has a solution. Please note that you do have to worry about zero pivots, though. Please note that you should not use the built-in function solve() to solve this system or use matrix inverses. The approach that you should employ is to construct an Upper Triangular Matrix
and then back-substitute to get the solution. Alternatively, you can augment the matrix A with vector b and jointly apply the Gauss Jordan elimination procedure.  
  
Please test it with the system below and it should produce a solution x = [-1.55, -0.32, 0.95]  
  
  
$\left[ \begin{matrix} 1 & 1 & 3 \\ 2 & -1 & 5 \\ -1 & -2 & 4 \end{matrix} \right] \begin{bmatrix} { x }_{ 1 } \\ { x }_{ 2 } \\ { x }_{ 3 } \end{bmatrix}\quad =\quad \begin{bmatrix} 1 \\ 2 \\ 6 \end{bmatrix}$    
  
*Given:*  
  
```{r}
# matrix A
A <- matrix(c(1,2,-1,1,-1,-2,3,5,4),3) 

# solution b
b <- c(1,2,6)    

```

```{r}
# function to solve for vector x in Ax = b.  Works only for 3x3 linear systems where there is a unique solution. 

matsolve33 <- function(mat,vec){
    if (!(is.numeric(mat) && is.numeric(vec))){
        warning("Both inputs must be numeric")
    }
    
    else if (!(is.matrix(mat) && nrow(mat) == 3 && ncol(mat) == 3)){
        warning("first input must be a 3x3 matrix")
    }
    else if (!(is.vector(vec) || (is.matrix(vec) && ncol(vec) == 1) || length(vec) == 3)){
        warning ("second input must be a vector of length 3 or a 3x1 matrix.")
    }
        
    else{
        # create augmented matrix [A b]
        aug <- cbind(mat,vec)  
          
        # permutation matrices for swapping rows
        p12 <- matrix(c(0,1,0,1,0,0,0,0,1),3, byrow=TRUE)
        p13 <- matrix(c(0,0,1,0,1,0,1,0,0),3,byrow=TRUE)
        p23 <- matrix(c(1,0,0,0,0,1,0,1,0),3,byrow=TRUE)
        
        # determine if row swapping is necessary for current row one
        # if element (1,1) is nonzero, do nothing
        # if element (1,1) is zero, check element (2,1).  if it is nonzero, swap rows 1 and 2
        # otherwise, swap rows 1 and 3
        
        which(c(1,2,3) == 3) != 1 

        if (which(aug[1,] != 0)[1] != 1){
        aug <- ifelse(which(aug[2,] != 0)[1] == 1, p12, p13) %*% aug
        }

        # perform row operations to set all entries below column one pivot equal to zero.

        for (i in 2:3){
            aug[i,] <- aug[i,] - aug[i,1]/aug[1,1] * aug[1,] 
        } 
        
        # if element (2,2) is not a pivot, then swap rows 2 and 3
        
        if (which(aug[2,] != 0)[1] != 2){
            aug <- p23 %*% aug
        }
        
        # perform row operations to set element (3,2) equal to zero 
        aug[3,] <- aug[3,] - aug[3,2] / aug[2,2] * aug[2,]

        
        # use back substitution to solve for x3, x2, and x1, respectively
        x3 <- aug[3,4] / aug[3,3]
        x2 <- (aug[2,4] - aug[2,3]*x3) / aug[2,2]
        x1 <- (aug[1,4] - aug[1,3]*x3 - aug[1,2]*x2) / aug[1,1]
    
        # return solution vector x: (x1, x2, x3)
        c(x1,x2,x3, use.names=FALSE)
        }
}

```

Let's verify that `matsolve33()` works using the given test data:
  
```{r}
matsolve33(A,b)
```
  
We can also double check our work using the built-in `solve()` function:  
  
```{r}
solve(A,b)

```
  


